---
title: "R Notebook"
output: html_notebook
---

Libraries used for this analysis

```{r}
require(ggplot2) # for data visualization
require(stringr) #extracting string patterns
require(Matrix) # matrix transformations
require(glmnet) # ridge, lasso & elastinet
require(xgboost) # gbm
require(randomForest)
require(Metrics) # rmse
require(dplyr) # load this in last so plyr doens't overlap it
require(caret) # one hot encoding
require(scales) # plotting $$
require(e1071) # skewness
require(corrplot) # correlation plot
```

Load both train and test data into dataframes

```{r}
train <- read.csv('data/house-prices-advanced-regression-techniques/train.csv', stringsAsFactors = FALSE)
test <- read.csv('data/house-prices-advanced-regression-techniques/test.csv', stringsAsFactors = FALSE)
```

## Data size and structure

```{r}
# combine the datasets
df.combined <- rbind(within(train, rm('Id','SalePrice')), within(test, rm('Id')))
dim(df.combined)
```
Our dataset is filled with many missing values, therefore, before we can build any predictive model we’ll clean our data by filling in all NA’s with appropriate values. For each column we’ll try to replace NA’s by using features that have a strong correlation, which will help us determine what values to fill in.

```{r}
na.cols <- which(colSums(is.na(df.combined)) > 0)
sort(colSums(sapply(df.combined[na.cols], is.na)), decreasing = TRUE)
```
```{r}
paste('There are', length(na.cols), 'columns with missing values')
```

```{r}
# helper function for plotting categoric data for easier data visualization
plot.categoric <- function(cols, df){
  for (col in cols) {
    order.cols <- names(sort(table(df.combined[,col]), decreasing = TRUE))
  
    num.plot <- qplot(df[,col]) +
      geom_bar(fill = 'cornflowerblue') +
      geom_text(aes(label = ..count..), stat='count', vjust=-0.5) +
      theme_minimal() +
      scale_y_continuous(limits = c(0,max(table(df[,col]))*1.1)) +
      scale_x_discrete(limits = order.cols) +
      xlab(col) +
      theme(axis.text.x = element_text(angle = 30, size=12))
  
    print(num.plot)
  }
}
```

# PoolQC: Pool Quality

This is the feature with the most missing values (2909). This feature represents the quality of the houses' pool (excellent, good, fair, etc.). We can reasonably assume that if this feature is NA that the house doesn't have a pool. However we can make use of the PoolArea feature to guarantee this, as if this is 0 then for sure the house doesn't have a pool.

```{r}
plot.categoric('PoolQC', df.combined)
```

```{r}
df.combined[(df.combined$PoolArea > 0) & is.na(df.combined$PoolQC),c('PoolQC','PoolArea')]
```
Of the 2909 rows, we see there are only 3 rows that fit this description (PoolArea > 0 and is.na = True).. In order to fill these gaps we can see what area (on average) corresponds to which pool quality, and then with that average we can set a PoolQC

```{r}
df.combined[,c('PoolQC','PoolArea')] %>%
  group_by(PoolQC) %>%
  summarise(mean = mean(PoolArea), counts = n()) 
```
Since only 10 houses have pools we can fill the rest of the NA values with 'NoPool'.. the 3 outliers we can fill in as so:

```{r}
df.combined[2421,'PoolQC'] = 'Ex'
df.combined[2504,'PoolQC'] = 'Ex'
df.combined[2600,'PoolQC'] = 'Fa'
df.combined$PoolQC[is.na(df.combined$PoolQC)] = 'None'
```

Now let's make sure it worked:

```{r}
df.combined %>%
  filter(is.na(PoolQC))
```

Next let's look into exploring more the Garage Related Variables:

GarageType: Garage location
GarageYrBlt: Year garage was built
GarageFinish: Interior finish of the garage
GarageCars: Size of garage in car capacity
GarageArea: Size of garage in square feet
GarageQual: Garage quality
GarageCond: Garage condition

